On the presentation of orthogonal groups over polynomial rings

### 2019-01-15 dl-paper.tex
Minor

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations

@@

\subsection{Relative Steinberg groups}
% Generation of relative Steinberg groups
Let us briefly remind the reader definitions and basic properties pertaining to relative Steinberg groups.
We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
%TODO: Define the category of pairs (so that all constructions below become functorial wrt it).

Thus, by our definition, the relative Steinberg group is a central extension of $\St(\Phi, R)$.
\begin{equation}\label{eq:relative-St}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r] & \St(\Phi, R, I) \ar[r] & \St(\Phi, R) \ar[r]^{\pi^*} & \St(\Phi, R/I) \ar[r] & 1 }
\end{equation}
It is not hard to show that $C(\Phi, R, I) = 0$ in the case when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see~\cite[Lemma~8]{S15}. In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. Thus we need to work with unstable $\K_3$-functors for which we choose the definition a la Quillen (cf.~\cite{Pa89})

\begin{df} Suppose that $\St(\Phi, R)$ is perfect and super-perfect (i.\,e. the homology groups $\HH_i(\St(\Phi, R), \ZZ)$ vanish for $i=1,2$.) Denote by $X$ the result of application of Quillen's $+$-construction to the discrete classifying space $B\G(\Phi, R)$ with respect to the elementary subgroup $\E(\Phi, R)$. Now define $\K_3(\Phi, R)$ as the third homotopy group $\pi_3(X)$. \end{df}
%TODO: Give reference on the definition of the plus-construction?

\begin{prop} The group $\K_3(\Phi, R)$ is isomorphic to $\HH_3(\St(\Phi, R), \ZZ)$. In particular, it does not depend on the isogeny form of $\G(\Phi, R)$ used to define it. Moreover, for a ring epimorphism $R \to R/I$ there is a functorial exact sequence
\[ \xymatrix{\K_3(\Phi, R) \ar[r] & \K_3(\Phi, R/I) \ar@{->>}[r] & C(\Phi, R, I).} \]
\end{prop}

>>>>

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{cond}[lemma]{Condition}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations

@@

\subsection{Relative Steinberg groups}

The aim of this section is to briefly remind the reader definitions and basic properties pertaining to relative Steinberg groups. We refer the reader to~\cite[Section~3]{S15} for a more detailed exposition.
%TODO: Define the category of pairs (so that all constructions below become functorial wrt it).

...

%Thus, by our definition, the relative Steinberg group is a central extension of $\St(\Phi, R)$.
\begin{equation}\label{eq:relative-St}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r] & \St(\Phi, R, I) \ar[r] & \St(\Phi, R) \ar[r]^{\pi^*} & \St(\Phi, R/I) \ar[r] & 1 }
\end{equation}
It is not hard to show that $C(\Phi, R, I) = 0$ in the special case when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see~\cite[Lemma~8]{S15}.

Unstable relative Steinberg groups share many good properties with their stable counterparts studied in~\cite{Lo78} provided the following assumptions on $\Phi$ hold: \begin{enumerate}
 \item $\St(\Phi, R)$ is perfect and super-perfect for all commutative $R$ (i.\,e. the homology groups $\HH_i(\St(\Phi, R), \ZZ)$ vanish for $i=1,2$).
 \item for all commutative rings $R$ the group $\K_2(\Phi, R)$ is central in $\St(\Phi, R)$.
\end{enumerate}
For example, under these assumptions it can be shown that the map $\St(\Phi, R, I) \to \St(\Phi, R)$ is a universal relative central extension of a pair $(\St(\Phi, R/I), \St(\Phi, R))$ in the sense of~\cite{Lo78} (the proof is identical to~\cite[Corollary~3]{S15}).

In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\begin{df}  
Denote by $X$ the topological space obtained from the (discrete) classifying space $B\G(\Phi, R)$ by means of Quillen's $+$-construction with respect to the elementary subgroup $\E(\Phi, R)$ (which is perfect thanks to the first assumption). Now set $\K_3(\Phi, R) = \pi_3(X)$. \end{df}
%TODO: Give reference on the definition of the plus-construction?

\begin{prop} The group $\K_3(\Phi, R)$ is isomorphic to $\HH_3(\St(\Phi, R), \ZZ)$. In particular, it does not depend on the isogeny form of $\G(\Phi, R)$ used to define it. Moreover, for a ring epimorphism $R \to R/I$ there is a functorial exact sequence
\[ \xymatrix{\K_3(\Phi, R) \ar[r] & \K_3(\Phi, R/I) \ar@{->>}[r] & C(\Phi, R, I).} \]
\end{prop}
\begin{proof}
 %TODO: First part.

 By~\cite[Th{\'e}or{\`e}me~2]{Lo78} the relative homology group $\HH_3(\St(\Phi, R/I), \St(\Phi, R), \ZZ)$ can be identified with the kernel $C(\Phi, R, I)$ of the relative central extension $\St(\Phi, R, I) \to \St(\Phi, R)$.
\end{proof}
